Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-07-12 Manuscript/sec2-review-of-graph-neural-network.tex
Update the section of reviewing GNN units.

\begin{figure}
	\centering
	\includegraphics[width=0.95\columnwidth]{figs/illustration/GNN_common_architecture.png}
	\caption{Structure of a typical graph neural network. (a) Demo GNN; (b) GNN unit; (c) Demo graph. The target application is the node classification. The demo GNN has two GNN layers.}
	\label{fig:general_structure_of_gnn}
\end{figure}

@@

A GNN layer consists of $n$ GNN units, where $n$ is the number of vertices in $\mathcal{G}$.
Each GNN unit correspond to a vertex in $\mathcal{G}$.
In the first GNN layer (Layer 0), the GNN unit of the vertex $v_i$ collects input feature vectors of the vertices $\boldsymbol{x}_j$ that are adjacent to $v_i$ in $\mathcal{G}$ (i.e., $j \in \mathcal{N}(v_i)$) from the input layer.
In other words, the connection between the input layer and the first GNN layer is determined by the topology of $\mathcal{G}$.
In the traditional neural networks, the neural units of neighboring layers are fully connected.
In GNNs, two GNN units are connected only if they have an edge between them in $\mathcal{G}$, which is usually \emph{sparse}.
After aggregating input feature vectors and applying non-linear transformation, the GNN unit outputs a hidden feature vector $\boldsymbol{h}^1_i$ for $v_i$.

In the next GNN layer (Layer 1), the GNN unit of $v_i$ collects the hidden feature vectors $\boldsymbol{h}^1_j$ of $v_i$'s neighbors (i.e., $j \in \mathcal{N}(v_i)$) from the \emph{previous} GNN layer.
The connection between neighboring GNN layers are also determined by $\mathcal{E}$ of $\mathcal{G}$.
Based on the collected hidden vectors, the GNN unit in Layer 1 outputs a new hidden feature vector $\boldsymbol{h}^2_i$ for $v_i$.
A GNN allows to stack more GNN layers to support deeper graph analysis.
%In \figurename~\ref{fig:general_structure_of_gnn}, we shows a GNN with two layers.

Assume there are $L$ GNN layers.
The last GNN layer (Layer $L-1$) outputs a hidden feature vector $\boldsymbol{h}^{L}_i$ for every vertex $v_i$.
As an embedding vector, $\boldsymbol{h}^L_i$ encodoes the knowledge learned from the input layer and all the previous GNN layers.
Finally, $\boldsymbol{h}^L_i$ is affected by $v_i$ and all the vertices in the $L$-hop neighborhood of $v_i$.
Analyzing with deeper GNNs means analyzing each vertices with a wider scope.

@@

>>>>

\begin{figure}
	\centering
	\includegraphics[width=1.1\columnwidth]{figs/illustration/GNN_common_architecture.png}
	\caption{Structure of a typical graph neural network. (a) Demo GNN, (b) Demo graph. The target application is the node classification. The demo GNN has two GNN layers.}
	\label{fig:general_structure_of_gnn}
\end{figure}

@@

A GNN layer consists of $n$ GNN units, where $n$ is the number of vertices in $\mathcal{G}$.
Each GNN unit correspond to a vertex in $\mathcal{G}$.
In the first GNN layer (Layer 0), the GNN unit of the vertex $v_i$ collects input feature vectors of itself and the vertices $\boldsymbol{x}_j$ that are adjacent to $v_i$ in $\mathcal{G}$ (i.e., $j \in \mathcal{N}(v_i)$) from the input layer.
After aggregating input feature vectors and applying non-linear transformation, the GNN unit outputs a hidden feature vector $\boldsymbol{h}^1_i$ for $v_i$.
Take the demo DNN in \figurename~\ref{fig:general_structure_of_gnn}(a) as the example.
Since $\mathcal{N}(v_3) = \{v_1, v_2, v_4, v_5, v_6\}$, the GNN unit of $v_1$ at layer 0 collects the feature vectors \{$\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{x}_3$, $\boldsymbol{x}_4$, $\boldsymbol{x}_5$, $\boldsymbol{x}_6$\} from the input layer and outputs $\boldsymbol{h}^1_1$.
We elaborate on the details of GNN units later.

The connection between the input layer and the first GNN layer is determined by the topology of $\mathcal{G}$.
In the traditional neural networks, the neural units of neighboring layers are fully connected.
In GNNs, two GNN units are connected only if they have an edge between them in $\mathcal{G}$.
Real-world graphs are usually very \emph{sparse}, i.e. $|\mathcal{E}| \ll |\mathcal{V}|^2$.

In the next GNN layer (Layer 1), the GNN unit of $v_i$ collects the hidden feature vectors of itself $\boldsymbol{h}^1_i$ and its neighbors ($\boldsymbol{h}^1_j$ with $j \in \mathcal{N}(v_i)$) from the \emph{previous} GNN layer.
Based on the collected hidden vectors, the GNN unit in Layer 1 outputs a new hidden feature vector $\boldsymbol{h}^2_i$ for $v_i$.
Though there are only two GNN layers in \figurename~\ref{fig:general_structure_of_gnn}, a GNN allows to stack more GNN layers to support deeper graph analysis.
%In \figurename~\ref{fig:general_structure_of_gnn}, we shows a GNN with two layers.

Assume there are $L$ GNN layers.
The last GNN layer (Layer $L-1$) outputs a hidden feature vector $\boldsymbol{h}^{L}_i$ for every vertex $v_i$.
As an embedding vector, $\boldsymbol{h}^L_i$ encodoes the knowledge learned from the input layer and all the previous GNN layers.
Since $\boldsymbol{h}^L_i$ is affected by $v_i$ and the vertices in the $L$-hop neighborhood of $v_i$, analyzing a graph with a \emph{deeper} GNN means analyzing each vertex with a \emph{wider} scope.

@@

GNN units are building blocks of a GNN.
A GNN layer consists of $|\mathcal{V}|$ units.
Each vertex has a corresponding GNN unit.
A GNN unit is a small neural network with the structure shown in \figurename~\ref{fig:general_structure_of_gnn}.
The GNN unit of $v_i$ at layer $l$ receives hidden feature vectors $\boldsymbol{h}^l_j$ from the GNN units of itself and its neighbors ($j \in \{i\} \cup \mathcal{N}(v_i)$) at the previous GNN layer \footnote{For the GNN layer 0, GNN units receive input feature vectors, i.e., $\boldsymbol{h}^0_i=\boldsymbol{x}_i$}.
The GNN unit aggregates the received hidden feature vectors, applies non-linear transformations, and outputs a new hidden feature vector $\boldsymbol{h}_i^{l+1}$.

\begin{figure}
	\centering
	\includegraphics[width=0.6\columnwidth]{figs/illustration/GNN_Unit.png}
	\caption{The GNN Unit of $v_3$ at the GNN layer $l$ with the graph $\mathcal{G}$ in \figurename~\ref{fig:general_structure_of_gnn}(b).}
	\label{fig:GNN_unit}
\end{figure}

We follow the message-passing model \cite{gilmer_messgae_passing} to formally define a GNN unit.
The message-passing model is widely used in the cutting-edge GNN training systems like PyTorch Geometric (PyG) \cite{PyG} and Deep Graph Library (DGL) \cite{DGL}.
\figurename~\ref{fig:GNN_unit} shows the structure of a GNN unit.
